---
title: "McAndrew Final Project JHU Course 3 R Notebook"
output: html_notebook
---
```{r}
install.packages("readr")
library(readr)
install.packages("dplyr")
library(dplyr)
install.packages("stringr")
library(stringr)
```{r}
Xtrain <- read.table("C:/Users/maure/Documents/R/UCI HAR Dataset/train/X_train.txt")
Ytrain <- read.table("C:/Users/maure/Documents/R/UCI HAR Dataset/train/Y_train.txt")
Subtrain <- read.table("C:/Users/maure/Documents/R/UCI HAR Dataset/train/subject_train.txt")
Xtest <- read.table("C:/Users/maure/Documents/R/UCI HAR Dataset/test/X_test.txt")
Yest <- read.table("C:/Users/maure/Documents/R/UCI HAR Dataset/test/Y_test.txt")
subtest <- read.table("C:/Users/maure/Documents/R/names(subtest)[names(subtest) == "V1"] <- "Subject ID"UCI HAR Dataset/test/subject_test.txt")
features <- read.table("C:/Users/maure/Documents/R/UCI HAR Dataset/features.txt")
activities <- read.table("C:/Users/maure/Documents/R/UCI HAR Dataset/activity_labels.txt")

```
#Combine test files
# First rename V1 columns in subtest and Ytest to avoid multiple columns with same name (V1) in combined dataset
```{r}names(subtest)[names(subtest) == "V1"] <- "Subject ID"
      names(Ytest)[names(Ytest) == "V1"] <- "Activity"
```
#Combine the three test datasets (subtest, Ytest, and Xtest) into a new dataset called Test.
```{r}
Test <- cbind(subtest, Ytest, Xtest)
```
#Combine train files
# First rename V1 columns in subtrain and Ytrain to avoid multiple columns with same name(V1) in combined dataset

```{r}
names(Subtrain)[names(Subtrain) == "V1"] <- "Subject ID"
names(Ytrain)[names(Ytrain) == "V1"] <- "Activity"
```
#Combinbe the three train datasets (Subtrain, Ytrain, and Xtrain) into a new
dataset called Train.

```
```{r}
Train <- cbind(Subtrain, Ytrain, Xtrain)
```
#Combine test and train datasets
```{r}
Testtrain <- rbind(Test, Train)
```
#Rename columns in Testtrain with Feature names
#First remove (temporarily) the Subject ID and Activity variables to create a new data set 
to rename all variables and call in Justvars.
```{r}
Justvars <- Testtrain %>% select(-`Subject ID`, Activity)
```
#Create a new dataset with Subject ID and Activity variables to merge after other variables are renamed with features
```{r}
Twovars <- Testtrain %>% select(`Subject ID`, Activity)
```
#Rename variables in Justvars with features file
```{r}
Justvars <- setNames(Justvars, features$V2)
```
#Combine Justvars and Twovars into new dataset called Testtrain1
```{r}
Testtrain1 <- cbind(Twovars, Justvars)
```
#Select only those columns for mean() and standard deviation or sd() and create a new dataset called meanstd
```{r}
meanstd <- Testtrain1 %>%
  select(SubjectID, Activity, contains("mean()") | contains("std()"))
```
#Add descriptive activity names to rename the numbers corresponding to activities in the Activity column. 
But first change class to factor and then create a vector of the activities for input into big dataset (Testtrain1).
```{r}
meanstd$Activity <- as.factor(meanstd$Activity)
newmeanstd <- str_replace_all(meanstd$Activity, c("1" = "walking", 
    "2" = "walkingupstairs", "3" = "walkingdownstairs","4" = "sitting",
    "5" = "standing", "6" = "laying"))
```
Replace numbered variable for Activity with descriptive names and create new dataset finalmeanstd.
First, create a new data frame (newact) from the new vector (newmeanstd) that contains the descriptive names of the numbered activities.
```{r}
newact <- as.data.frame(newmeanstd)
```
Change (mutate) the Activity column with the descriptive names in the main dataset that is renamed finalmeanstd

```{r}
finalmeanstd <- meanstd %>% mutate(Activity = newact$newmeanstd)
```
# Replace variable abbreviations (t, f) to represent (time, freq) and eliminated unneeded ().
```{r}
names(finalmeanstd) <- gsub(x = names(finalmeanstd), pattern = "^t", replacement = "time")
names(finalmeanstd) <- gsub(x = names(finalmeanstd), pattern = "^f", replacement = "freq")
names(finalmeanstd) <- gsub(x = names(finalmeanstd), pattern = "\\(", replacement = "")
names(finalmeanstd) <- gsub(x = names(finalmeanstd), pattern = "\\)", replacement = "")
```
#For each of the 30 subjects and for each of the 6 activities, calculate the mean of each unique measure 
of the mean() and std() variables. Save in a dataset called md
```{r}
md <- finalmeanstd %>%
+     group_by(`Subject ID`, Activity) %>%
+     summarize_all(list(mean))
```
# Save tidied dataset (md) to text file.
```{r}
write.table(md, file = "C:/Users/maure/Documents/R/JHU March/tidied.txt")
```




